---
title: "Rio 2016 Olympics"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
    theme: cerulean
runtime: shiny
resource_files:
- Rio_2016_Olympics.png
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(tidyverse)
library(packcircles)
library(countrycode)
library(ggiraph)
library(plotly)
library(lubridate)
library(eeptools)
library(maps)
library(mapproj)
library(DT)

dat <- read_csv("Olympics_2016_Rio_Athletes.csv")

dat <- dat %>%
  mutate(country = case_when(
    nationality == 'IOA' ~ 'Independent Olympians',
    nationality == 'ISV' ~ 'United States Virgin Islands',
    nationality == 'LIB' ~ 'Lebanon',
    nationality == 'ROT' ~ 'Refugee Olympic Team',
    nationality == 'SIN' ~ 'Singapore',
    nationality >0 ~ countrycode(nationality, "ioc", "country.name"))) %>%
  mutate(continent = case_when(
    nationality == 'IOA' ~ 'Independent',
    nationality == 'ISV' ~ 'Americas',
    nationality == 'LIB' ~ 'Asia',
    nationality == 'ROT' ~ 'Independent',
    nationality == 'SIN' ~ 'Asia',
    nationality == 'KOS' ~ 'Europe',
    nationality >0 ~ countrycode(nationality, "ioc", "continent"))) %>%
  mutate(continent_id = case_when(
    continent == 'Americas' ~ 1,
    continent == 'Europe' ~ 2,
    continent == 'Asia' ~ 3,
    continent == 'Africa' ~ 4,
    continent == 'Oceania' ~ 5,
    continent ==  'Independent' ~ 6)) %>%
  mutate(total = gold + silver + bronze) %>%
  mutate(dob = mdy(dob))

cent <- function(x, year=1940){
  m <- year(x) %% 100
  year(x) <- ifelse(m > year %% 100, 1900+m, 2000+m)
  x
}

dob2 = as.tibble(cent(dat$dob))
dat <- add_column(dat,as.vector(dob2)) %>%
  rename("dob2"="value") %>% 
  mutate(age = floor(interval(start= dob2, end=mdy("08-05-2016"))/                      
                     duration(n=1, unit="years")))

```

The athletes {data-orientation=rows}
======================================

Row {data-height=500}
-----------------------------------------------------------------------

### Chart A

```{r, echo=FALSE, warning=FALSE}

fig_dat1 <- dat %>% select(nationality,country,continent,continent_id)

temp_dat1 <- dat %>% select(country)
temp_dat1 <- as_tibble(table(temp_dat1)) %>%
  rename("country"="temp_dat1") %>% 
  rename("number_of_athletes"="n")

fig_dat1 <- fig_dat1 %>% distinct(country, .keep_all = TRUE)

fig_dat1 <- left_join(temp_dat1,fig_dat1,by="country") %>%
  arrange(desc(number_of_athletes))

packing <- circleProgressiveLayout(fig_dat1$number_of_athletes,sizetype='area')
fig_dat1 <- add_column(fig_dat1,packing)

dat.gg <- circleLayoutVertices(packing,npoints=100)

continent_id <- rep(fig_dat1$continent_id,each=101)
dat.gg <- add_column(dat.gg,continent_id)

number_of_athletes <- rep(fig_dat1$number_of_athletes,each=101)
dat.gg <- add_column(dat.gg,number_of_athletes)

country <- rep(fig_dat1$country,each=101)
dat.gg <- add_column(dat.gg,country)

nationality <- rep(fig_dat1$nationality,each=101)
dat.gg <- add_column(dat.gg,nationality)

if (requireNamespace("ggiraph")) {
  
gg1 <- ggplot(data=dat.gg) +
      ggiraph::geom_polygon_interactive(
               aes(x=x,
                   y=y,
                   group=id,
                   fill=as.factor(continent_id),
                   tooltip = c(paste0(country, " Athletes: \n",number_of_athletes)),data_id=x))+
  scale_fill_manual(
    labels = c(
      "Americas",
      "Europe",
      "Asia",
      "Africa",
      "Oceania",
      "Independent"),
    values=alpha(c(
      rgb(214,39,40,max=255),
      rgb(44,160,44,max=255),
      rgb(31,119,180,max=255),
      rgb(140,86,75,max=255),
      rgb(227,119,194,max=255),
      rgb(255,127,14,max=255)),
      0.75)) +
  geom_text(data=filter(fig_dat1,number_of_athletes > 100),
            aes(x=x,
                y=y,
                size=number_of_athletes,
                label=nationality),color="white") + 
  scale_size(range = c(5, 9)) +
  guides(alpha=FALSE,
         size=FALSE,
         color = guide_legend(override.aes = list(alpha = 0.1)),
         fill=guide_legend(title="Region",override.aes = list(size=10))) +
  labs(title="Number of athletes by nationality & region") +
  theme(legend.position = c("right"),
        legend.title = element_text(face = "bold",size=20),
        legend.text = element_text(size = 20),
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        plot.title = element_text(hjust = 0.5,face="bold",size=36)) +
  coord_equal()

 renderGirafe({ggiraph::ggiraph(ggobj = gg1, width_svg = 20, height_svg = 10)})#, options = list(
   #opts_selection(type = "multiple", css = "fill:blue;stroke:gray;r:5pt;")))})
  
}

```

### Chart B

```{r, echo=FALSE, warning=FALSE}

my_world_map <- map_data("world") %>%
  rename("country"="region")

my_world_map <- my_world_map %>%
  mutate(nationality = countrycode(country, "country.name", "ioc"))

fig_dat2 <- dat %>% select(nationality,country)

temp_dat2 <- dat %>% select(country)
temp_dat2 <- as_tibble(table(temp_dat2)) %>%
  rename("country"="temp_dat2") %>% 
  rename("Athletes"="n")

fig_dat2 <- fig_dat2 %>% distinct(country, .keep_all = TRUE)

fig_dat2 <- inner_join(temp_dat2,fig_dat2,by="country") %>%
  arrange(desc(Athletes))

my_dat2_combined <- left_join(my_world_map,fig_dat2,by="nationality")

gg2 <- ggplot()+
  geom_polygon(data=my_dat2_combined, mapping=aes(x=long,y=lat,group=group,text = paste(country.x), fill= Athletes,colour="gray10"))+
  scale_fill_distiller(palette=7,direction=1)+
  scale_colour_manual(values = c("gray10"))+
  theme_bw()+ 
  guides(color=FALSE)+
  coord_map(xlim=c(-180,180))+
  labs(title="Global distribution of athletes",x="",y="")+
  theme(legend.position = c("right"),
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),
        plot.title = element_text(size=16,face="bold"))

ggplotly(gg2,tooltip=c("text","fill")) %>% 
  hide_legend()

# ggplotly(gg2,width = 1000, height = 480,tooltip=c("text","fill")) %>% 
#   hide_legend()

```

Row {data-height=430}
-----------------------------------------------------------------------

### Contents

#### **Page 1:** explores the range of athletes that competed in the Rio 2016 Olympic Games. Click and drag on the map to zoom on regions, double click to reset.  

<br>

#### **Page 2:** explores the medals and medalists of the Games. Configure the medals, nationality & sport input parameters on the sidebar to filter the underlying data.  

<br>

#### **Page 3:** explores the attributes of medalists. What are observed common attributes of Olympic medalists? Configure sport & sex input parameters on the sidebar to filter the underlying data.  

<br>

#### **Page 4:** shows the total medals table. Please note that this counts all individual medals awarded, whereas the official ranking tallies only 1 medal for a team placing in a sport.  

```{r}
library(knitr)
renderImage(
#list(src = "C:/Users/Matthew/Desktop/Johns Hopkins Data Visualization & Dashboarding/5 - Data Visualization Capstone/Rio_2016_Olympics.png",contentType = 'image/png',
  list(src = "Rio_2016_Olympics.png",contentType = 'image/png',
         width = 100,
         height = 100,
     deleteFile=FALSE)
)
```

### Chart C

```{r, echo=FALSE}
#pie chart of gender, filter by country

fig_dat3 <- dat %>% 
  #filter(nationality == USA) %>%
  select(sex) 
fig_dat3 <- as_tibble(table(fig_dat3)) %>%
  rename("count"="n") %>%
  rename("sex"="fig_dat3") %>%
  mutate(per=count/sum(count)) %>%
  arrange(count)

plot_ly(fig_dat3, labels = ~sex, values = ~per, type = 'pie', textinfo = 'label+percent', 
        insidetextfont = list(size = 24), hoverinfo = 'text', text = ~paste(sex,"athletes:",count),
        marker = list(colors = c("rgba(214,39,40,0.75)","rgba(31,119,180,0.75"),line = list(color = '#000000', width = 2))) %>%
  layout(title = "<b>Percentage of athletes by gender<b>",font=list(size = 12),
         legend=list(title=list(text='<b> Sex </b>')))

```

The medalists {data-orientation=rows}
======================================

Input {.sidebar}
-----------------------------------------------------------------------

### Input parameters

```{r}

sidebar_dat1 <- dat %>%
  filter(total > 0)

selectInput(inputId="my_medal",label="medals",choices=c("total","gold","silver","bronze"),multiple=FALSE,selected="total")

selectInput(inputId="my_nationality",label="nationality",choices=sort(unique(sidebar_dat1$nationality)),multiple=TRUE,selected=unique(sidebar_dat1$nationality))

selectInput(inputId="my_sport",label="sport",choices=sort(unique(sidebar_dat1$sport)),multiple=TRUE,selected=unique(sidebar_dat1$sport))

```


Row {data-height=500}
-----------------------------------------------------------------------

### Chart D

```{r, echo=FALSE}

renderGirafe({
  
fig_dat4 <- dat %>% select(nationality,sport,country,sex,!!as.symbol(input$my_medal)) %>%
  rename("medal"=input$my_medal) %>%
  filter(nationality %in% input$my_nationality) %>%
  filter(sport %in% input$my_sport) %>%
  group_by(country,sex) %>% 
  mutate(count = n())

fig_dat4 <- aggregate(cbind(medal) ~ country + sex + nationality, data = fig_dat4, FUN = sum, na.rm = TRUE) %>%
  filter(medal >0 ) %>% #need to filter to remove zeros for x & y of dat.gg
  arrange(country)

packing <- circleProgressiveLayout(fig_dat4$medal,sizetype='area')
fig_dat4 <- add_column(fig_dat4,packing)

dat.gg <- circleLayoutVertices(packing,npoints=100)

sex <- rep(fig_dat4$sex,each=101)
dat.gg <- add_column(dat.gg,sex)

medal <- rep(fig_dat4$medal,each=101)
dat.gg <- add_column(dat.gg,medal)

country <- rep(fig_dat4$country,each=101)
dat.gg <- add_column(dat.gg,country)

nationality <- rep(fig_dat4$nationality,each=101)
dat.gg <- add_column(dat.gg,nationality)

max(dat.gg$medal, na.rm = TRUE)

if(max(dat.gg$medal, na.rm = TRUE) <= 5){
  display_filter <- 0
} else if(max(dat.gg$medal, na.rm = TRUE) <= 10){
  display_filter <- 5
} else if(max(dat.gg$medal, na.rm = TRUE) <= 15){
  display_filter <- 10
} else {
  display_filter <- 15
}

if (requireNamespace("ggiraph")) {
  
gg4 <- ggplot(data=dat.gg) +
      ggiraph::geom_polygon_interactive(
               aes(x=x,
                   y=y,
                   group=id,
                   fill=as.factor(sex),
                   tooltip = c(paste0(country, " ", sex, " ", input$my_medal, " Medalists: \n",medal)),data_id=x))+
  scale_fill_manual(
    labels = c(
      "Female",
      "Male"),
    values=alpha(c(rgb(214,39,40,max=255),rgb(31,119,180,max=255)),0.75)) +
  geom_text(data=filter(fig_dat4,medal > display_filter),
            aes(x=x,
                y=y,
                size=medal,
                label=nationality),color="white") + 
  scale_size(range = c(5,9)) +
  guides(alpha=FALSE,
         size=FALSE,
         color = guide_legend(override.aes = list(alpha = 0.1)),
         fill=guide_legend(title="Sex")) +
  labs(title="Medals by nationality and sex") +
  theme(legend.position = c("right"),
        legend.title = element_text(face = "bold",size=22),
        legend.text = element_text(size = 22),
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        plot.title = element_text(hjust = 0.5,size=36,face="bold")) +
  coord_equal() #+
  #ylim(-20,20)

#ggiraph::ggiraph(ggobj = gg4)
  
 ggiraph::ggiraph(ggobj = gg4, width_svg = 16, height_svg = 10, width = 1)
  
}
})

```

### Chart E

```{r, echo=FALSE}
#height vs weight, medals=size of points, facet gender?

renderPlotly({
  
gg5 <- (fig_dat5 <- dat %>% select(name,nationality,country,sport,sex,height,weight,!!as.symbol(input$my_medal)) %>%
  rename("medal"=input$my_medal) %>%
  filter(nationality %in% input$my_nationality) %>%
  filter(sport %in% input$my_sport) %>%
  filter(medal > 0) %>%
  ggplot(
    aes(x=weight,y=height,fill=sex,size=medal,text = paste("Name:", name, "\nSex:", sex, "\nCountry:", country, "\nHeight:", height, "m\nWeight:", weight, "kg\n",input$my_medal,"medals:", medal))) +
    geom_jitter() +
    labs(title="<b> Medalists by height vs. weight & sex <b>",x="Weight (kg)",y="Height (m)") +
    scale_fill_manual(values=alpha(c(rgb(214,39,40,max=255),rgb(31,119,180,max=255)),0.5),name="")) +
    scale_size_continuous(range=c(3,10),name="") +
  theme_light() +
  ylim(1.4,2.2) +
  xlim(30,160)

ggplotly(gg5,tooltip=c("text")) %>%
  layout(legend = list(title = list(text = "<b>Sex</b>")))

})

#add slider for medals/size
  
```

Row {data-height=400}
-----------------------------------------------------------------------

### Chart F

```{r, echo=FALSE}

renderPlotly({
  
 fig_dat6 <- dat %>% select(nationality,sport,gold,silver,bronze,total) %>%
    filter(nationality %in% input$my_nationality) %>%
    filter(sport %in% input$my_sport) %>%
    group_by(sport) %>% 
    mutate(count = n())

  fig_dat6 <- aggregate(cbind(gold,silver,bronze,total) ~ sport, data = fig_dat6, FUN = sum, na.rm = TRUE) %>% 
    arrange(desc(total)) %>%
    pivot_longer(cols=c(gold,silver,bronze),names_to="medal",values_to="count")

fig_dat6$medal <- factor(fig_dat6$medal, levels = c("gold", "silver", "bronze"))

 if(input$my_medal != "total"){
   fig_dat6 <- fig_dat6 %>%
     filter(medal == input$my_medal)
 } 

if(input$my_medal == "gold"){
  my_fill <- c(alpha("gold"))
} else if(input$my_medal == "silver"){
  my_fill <- c(alpha("grey"))
} else if(input$my_medal == "bronze"){
  my_fill <- c(alpha(rgb(0.804,0.498,0.196)))
} else {
  my_fill <- c(alpha("gold"),alpha("grey"),alpha(rgb(0.804,0.498,0.196)))
}

gg6 <- ggplot(fig_dat6,aes(x=reorder(sport,-count),y=count,fill= medal,text = paste(count, medal, "medalists"),color="gray50"))+geom_bar(stat="identity")+labs(x="Sport",y="Medals",title="<b> Medals awarded by sport <b>")+
  scale_fill_manual(
    values=my_fill) + 
  scale_colour_manual(values = c("gray50"),name="")+
  theme_bw()+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1, color="darkgrey" ),
        axis.text.x = element_text(size=6))

ggplotly(gg6, tooltip=c("text")) %>% 
  hide_legend()

# ggplotly(gg6, width = 1750, height = 400, tooltip=c("text")) %>% 
#   hide_legend()

})

```

Attributes {data-orientation=rows}
======================================

Input {.sidebar}
-----------------------------------------------------------------------

### Input parameters

```{r}

sidebar_dat2 <- dat %>%
  filter(total > 0)

selectInput(inputId="my_sex",label="sex",choices=c("female","male"),multiple=TRUE,selected=c("female","male"))

selectInput(inputId="my_sport2",label="sport",choices=sort(unique(sidebar_dat2$sport)),multiple=TRUE,selected=unique(sidebar_dat2$sport))

```

Row {data-height=500}
-----------------------------------------------------------------------

### Chart G

```{r, echo=FALSE}

renderPlotly({
  
fig_dat7 <- dat %>% 
  select(sex, sport, weight, height, age, total) %>%
  filter(sex %in% input$my_sex) %>%
  filter(sport %in% input$my_sport2) %>%
  rename("weight_kg"="weight") %>% 
  rename("height_m"="height") %>%
  rename("age_years"="age")

fig_dat7 <- fig_dat7 %>%
  pivot_longer(cols=c(weight_kg,height_m, age_years),names_to="variable",values_to="metric")

fig_dat7 <- aggregate(total ~ metric + variable, data = fig_dat7, FUN = sum, na.rm = TRUE) 

fig_dat7$variable <- factor(fig_dat7$variable, levels = c("weight_kg", "height_m", "age_years"))

gg7 <-  ggplot(fig_dat7,
    aes(x=metric,y=total,fill=variable,color="black",group=1,text=paste0(variable, ": ",metric, "\nmedalists: ", total))) +
    geom_area() +
  geom_point(aes(x=metric, y=total)) +
    labs(title="<b> Distribution of medalists by weight, height & age <b>",x="",y="Medals") +
    facet_wrap(~variable,scales = "free") +
    scale_colour_manual(values = c("black"),name="") +
    scale_fill_manual(values = c(rgb(255,187,120,max=255),rgb(152,223,138,max=255),rgb(174,199,232,max=255)))+
  theme_light() +
  theme(plot.title = element_text(size=20,face="bold"),
        strip.text.x = element_text(size = 16, face = "bold"),
        strip.background = element_rect(color = "black", size = 1))

ggplotly(gg7, tooltip=c("text")) %>% 
  hide_legend() %>%
  layout(margin=list(t = 100))

})
#add filter by sex
#labels with unit

```

Row {data-height=500}
-----------------------------------------------------------------------

### Chart H

```{r, echo=FALSE}

renderPlotly({
  
fig_dat8 <- dat %>% 
  select(sex, sport, weight, height, age, total) %>%
  filter(sex %in% input$my_sex) %>%
  filter(sport %in% input$my_sport2) %>%
  filter(total > 0) %>%
  rename("weight_kg"="weight") %>% 
  rename("height_m"="height") %>%
  rename("age_years"="age")

fig_dat8 <- fig_dat8 %>%
  pivot_longer(cols=c(weight_kg,height_m, age_years),names_to="variable",values_to="metric")

fig_dat8$variable <- factor(fig_dat8$variable, levels = c("weight_kg", "height_m", "age_years"))

gg8 <- ggplot(fig_dat8,aes(x=as.factor(variable),y=metric,fill=variable)) +
  geom_boxplot() +
  facet_wrap(~variable,scales = "free") +
  labs(title="<b> Spread of medalists by weight, height & age <b>",x="",y="") +
  scale_fill_manual(values = c(rgb(255,187,120,max=255),rgb(152,223,138,max=255),rgb(174,199,232,max=255))) +
  theme_light() +
  theme(axis.text.x=element_blank(),
        plot.title = element_text(size=20,face="bold"),
        strip.text.x = element_text(size = 16, face = "bold"),
        strip.background = element_rect(color = "black", size = 1))

ggplotly(gg8) %>%
  hide_legend() %>%
  layout(margin=list(t = 100))

})

#tooltip custom for outliers, solution in shiny
#add filter by sex

```

Medals Table {data-orientation=rows}
======================================

Row {}
-----------------------------------------------------------------------

### Medals Table

```{r}

renderDataTable({
  
  fig_dat9 <- dat %>% 
  select(country, gold, silver, bronze, total)
  
  fig_dat9 <- aggregate(cbind(gold,silver,bronze,total) ~ country, data = fig_dat9, FUN = sum, na.rm = TRUE) %>% 
  arrange(desc(total))
  
  fig_dat9
    
})

```

