---
title: "Spatial Figures Peer Review"
output: html_document
---
# Matthew Pinkerton

## Problem 1

Create the following figure, using the data included in the R Markdown file.

```{r problem_1_data,include=FALSE}

library(tidyverse)
library(maps)
library(mapproj)

####make some data for painting the map

my_world_map <- map_data("world")

countries <- unique(my_world_map$region)

set.seed(987)
some_data_values <- data.frame(

  "region"=countries,
  
  "Score"=runif(252,0,100))

```

```{r problem_1,echo=FALSE,message=FALSE}

my_data_combined <- left_join(my_world_map,some_data_values,by="region")
ggplot(data=my_data_combined,mapping=aes(x=long,y=lat,group=group,fill=Score))+geom_polygon()+scale_fill_distiller(palette=5)

```

## Problem 2

Create the following figure, using the data included in the R Markdown file.


```{r problem_2_data,include=FALSE}

#####DO NOT MODIFY THIS CHUNK
set.seed(15)
Measurement<-rnorm(32,50,1)
```


```{r solution_p2,echo=FALSE,message=FALSE}

####Make sure you load any necessary libraries

####HINT: Use a filter command to get just maps of Costa Rica, Panama, and Nicaragua

####Hint: Use a filter command to put in points only for cities with a population of greater than 40,000. This should leave you with 32 cities.

####Hint: Use add_column to attach the "Measurement" variable to your data, and set that to the color aesthetic when you draw the points.

####Hint: set the colors for the city points with scale_color_distiller(palette=7)

####Hint: set the size of all points to the value 5

CRPA <- my_world_map %>% filter(region =="Costa Rica" | region =="Panama" | region =="Nicaragua")

data2 <- world.cities %>% filter(country.etc =="Costa Rica" | country.etc == "Panama" | country.etc == "Nicaragua")

CRPA2 <- data2 %>% filter(pop > 40000) %>% add_column(Measurement)

ggplot(data=CRPA,mapping=aes(x=long,y=lat,group=group))+geom_polygon(fill="white",color="black")+geom_point(data=CRPA2,aes(x=long,y=lat,group=NULL,color=Measurement,size=5))+scale_color_distiller(palette=7)+guides(size=FALSE)

```

## Problem 3

Create the following figure, using the data included in the R Markdown file.

Note that the code in the .rmd file will import a set of simple features data for South America. Make sure you install any necessary packages. 

```{r problem_3_data,include=FALSE}

#####INSTALL THESE PACKAGES IF NECESSARY
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)

####DO NOT MODIFY
s_america<-ne_countries(scale="medium",continent='south america',returnclass="sf")

```

```{r problem3,echo=FALSE}

####Make sure you load any necessary libraries

####HINT: the s_america object created in the chunk above is a simple features object with many columns. Identify the correct column based on the solution figure and use it to color in the choropleth.

####HINT: Use scale_fill_distiller and palette 10 to the match the colors.

ggplot()+ geom_sf(data=s_america,aes(fill=pop_est))+scale_fill_distiller(palette=10)

```
