---
title: "Week 3 Peer Review"
output: html_document
---

```{r setup, include=FALSE}

######DO NOT MODIFY. This will load required packages and data.

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggthemes)

cces <- drop_na(read_csv(url("https://www.dropbox.com/s/ahmt12y39unicd2/cces_sample_coursera.csv?raw=1")))

cel <- drop_na(read_csv(url("https://www.dropbox.com/s/4ebgnkdhhxo5rac/cel_volden_wiseman%20_coursera.csv?raw=1")))

```

Following the directions on the Coursera assignment page, you will make four original visualizations. Note that the data for the CCES and CEL data are imported in code in the R Markdown file.

### Matthew Pinkerton

## Exercise 1

### Scatter plot

Is there correlation between legislative effective score and bills passed? The graphs below show a strong linear association between these 2 variables. Bills passed is likely a strong factor in determining LES. BY splitting by gender, we can see the same association apparent with both genders, with less females passing a large number of bills and scoring highly in LES.

```{r exercise1,echo=FALSE}

my_dat <- cel %>% 
  rename("gender"="female") %>% 
  mutate(gender=recode(gender,'0'='Male','1'='Female'))

ggplot(my_dat,aes(x=les,y=all_pass,color=gender))+geom_jitter()+labs(x="legislative effective score",y="Bills passed",title="Plot of LES vs. Bills passed")+scale_color_manual(values=c("red","blue"))+facet_wrap(~gender)+geom_smooth(method=lm)

```

## Exercise 2

### Histogram (distribution plot)

How many MCs are in the member's state delegation? The below histogram shows a distribution of delegation size. We see right skew with most delegations having fewer numbers of MCs, with a small peak in the largest membership bin.

```{r exercise2,echo=FALSE}

ggplot(my_dat,aes(x=deleg_size))+geom_histogram(bins=10,color="black",fill="light blue")+labs(title="Distribution of delegation size",x="Number of MCs",y="Frequency")+theme_bw()
  
```

## Exercise 3

What is the highest level of education completed within the cces sample? The bar graph below tells us that it is "4-year".

```{r exercise3,echo=FALSE}

my_dat <- cces %>% mutate(educ=recode(educ,'1'='None','2'='High school', '3' = 'Some college', '4' = '2-year', '5' = '4-year','6' = 'Post-grad'))
ggplot(my_dat,aes(x=educ, fill=educ))+geom_bar()+labs(x="Education",y="Count") + guides(fill=FALSE)+labs(title="Level of Education")

```

## Exercise 4

Which month recorded the highest and lowest temperatures in the dataset? The line graph indicates that the lowest month was February and the highest month was August.

```{r exercise4,echo=FALSE}

month <- c(1,2,3,4,5,6,7,8,9,10,11,12)
average_temperature <- c(5,4,5,9,17,22,22,23,20,16,12,8)

my_data <- tibble(month,average_temperature)

ggplot(my_data,(aes(x=month,y=average_temperature)))+geom_line()+ guides(fill=FALSE)+labs(title="Plot of average temperature",y="Average temperature (degrees celcius)")+scale_x_continuous(breaks = seq(0, 12, 1))+annotate(geom = "text", x = 2, y = 4, label = "Lowest:", vjust = -2.5)+annotate(geom = "text", x = 2, y = 4, label = "February", vjust = -1.3)+annotate(geom = "text", x = 8, y = 23, label = "Highest:", vjust = 3)+annotate(geom = "text", x = 8, y = 23, label = "August", vjust = 4.2)+theme_solarized()

```
